Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-04 doc/sc2012/circuit.tex
revision

Listing~\ref{lst:code_ex} shows code for an electrical
circuit simulation, which illustrates the core features of the Legion programming model.
The circuit simulation takes a collection of
of wires and nodes where wires meet.  
The simulation repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.  We first
describe the partitioning of regions in Section~\ref{subsec:partitioning}
and then describe the data type declarations in Section~\ref{subsec:datatypes}.

\subsection{Regions and Partitions}
\label{subsec:partitioning}

Line 9 declares the main simulator function, which takes a {\tt
Circuit} to be simulated.  This function specifies the regions it 
accesses and the privileges and coherence it requires of those regions.
In this case, the {\tt RWE} annotation specifies that the function
accesses the regions {\tt c.r\_all\_nodes} and {\tt c.r\_all\_wires}
with read-write privileges and {\em exclusive} coherence (i.e., no other
function can access these two regions concurrently).  The simulation 
reads and writes all nodes and wires, and it must be done
with exclusive access to ensure correct results.  Privileges specify what
the function can do with the regions; coherence specifies what other
functions can do with the regions concurrently.  Functions that
declare their accessed regions, privileges, and coherence are called {\em tasks}
and are the unit of parallel execution in Legion.

Lines 15-21 are responsible for partitioning the circuit into {\tt MAX\_PIECES}
pieces that can be worked on in parallel.  To partition a region, we
provide a {\em coloring}, which is a relation between the elements
of a region and a set of colors.  A {\em partition} is an object which given
a coloring and a region $r$ contains for each color $c$ a {\em subregion} of $r$
containing all the elements of $r$ colored $c$.

Wires are simply partitioned in {\tt MAX\_PIECES} subregions (line 16).  
The nodes are first partitioned into 
{\em private} and {\em shared} nodes (line 21).  Private nodes
are touched by a single task in a phase, while shared nodes may
be referenced by multiple tasks in a phase.  The thick lines in Figure~\ref{sfig:part_fig:pvs}
show how the nodes in a small graph might be partitioned into three subsets.  Orange
nodes are shared and blue nodes are private.  Note that
this partition is disjoint since each node has one color.
%Our current implementation uses small
%integers or booleans for the names of colors, but any enumerable set would be
%fine.  The coloring relation need not be total - it can leave out some elements entirely.  And although it is not required to be injective (it can map a single
%element to multiple colors), the fact that injective mappings will always
%result in disjoint subregions makes them strongly encouraged (when possible).
%Our example uses four colorings to create circuit partitions.

We next further partition the private and shared subregions into {\tt
  MAX\_PIECES} disjoint subregions each (lines 20-23); the private and shared node subregions
for the upper-left patch of the graph is shown in
Figures~\ref{sfig:part_fig:p_i} and \ref{sfig:part_fig:s_i}
respectively.

We partition the shared nodes a second time using a
different coloring into the sets of ghost nodes for each task.  This 
partition is noteworthy because the subregions are not disjoint:
ghost nodes may be read by multiple tasks, and thus a node $n$  may belong 
to more than one ghost region, which corresponds to $n$ being assigned
more than one color by the coloring {\tt node\_neighbor\_map}.
The ghost subregion for the upper left patch of the graph
 is shown in Figure~\ref{sfig:part_fig:g_i}.

Figure~\ref{sfig:part_fig:tree} shows the final hierarchy of node
partitions;  the $*$ symbol indicates a partition is disjoint. This
{\em region tree} data structure plays an important role in scheduling
tasks for out-of-order execution (see Section~\ref{sec:soop}).

%(These subsets are hopefully reasonably compact, but the correctness
%of the simulation is not dependent on that.)  
%Once the subsets are known, the \emph{node\_owner\_map} is created by assigning
%each node the color corresponding to its subset.  The 
%\emph{wire\_owner\_map} assigns each wire to the same subset as its ``in\_node''.
%The \emph{node\_nghbr\_map} maps a node to color(s) of all wires that 
%connect to it.  Finally, the \emph{node\_sharing\_map} is derived from the
%\emph{node\_nghbr\_map}, with a node colored ``true'' if any colors other than 
%its own were used, and ``false'' if the only wires that connect to a node are
%in the same piece.
%Figure~\ref{fig:part_fig} shows how the partitions are defined.

%Line 19 uses the \emph{wire\_owner\_map} to partition the wires region of the
%circuit into a subregion for each piece, but the partitioning of the nodes is
%more complicated due to the sharing that is necessary between the pieces.
%First, the \emph{node\_sharing\_map} is used to create two subregions: $p\_nodes\_pvs[false]$ contains all the nodes that are private to some piece (i.e. will
%never be needed for the computations in any other piece), while $p\_nodes\_pvs[true]$ contains nodes that will be accessed by multiple pieces' computations.
%Each of these subregions is then partitioned using the $node\_owner\_map$ to
%create the subregions owned by each computation.  These are the $p_i$ and $s_i$
%subregions, shown in Figures \ref{sfig:part_fig:p_i} and \ref{sfig:part_fig:s_i}.  Finally, the $g_i$ subregions (shown in Figure~\ref{sfig:part_fig:g_i}) are
%created using the $node\_nghbr\_map$ to 
%create subregions that include the ``ghost'' nodes needed to perform each
%piece's calculations.

%With the partitioning operations completed, the various subregions are
%recorded in an array of {\tt CircuitPiece} structures.  Lines 32-33 fill in multiple
%fields of a {\tt CircuitPiece} at once.  The need for such a statement will be described in detail in 
%section \ref{subsec:datatypes}.

%Because a region relationship can have
%fields with types that refer its own fields' values, it is often not possible
%to assign fields one at a time without violating the type checking rules.
%The simultaneous assignment operator asks the type checker only to make sure
%that the fields in the structure would have the right (self-referential) types
%after all the fields have been changed.

Lines 31-57 form the bulk of the actual simulation, performing three
passes over the circuit in each iteration.  Each pass loops over an
array of pieces (constructed on lines 27-30 from the partitions),
spawning a task for each piece.  There are no explicit requests for
parallel execution (the default semantics of {\tt spawn} is sequential
execution of the spawned tasks) nor is there synchronization required
between the passes.  Both the fact that the tasks can be run in
parallel within a pass and the required inter-pass dependencies are
determined automatically by the Legion runtime based on the region
access annotations on the task declarations.  The tasks spawned on
lines 32-34 are {\em subtasks} of the main {\tt simulate\_circuit}
task. A subtask can only take a region argument that is a region argument of the
parent task or a subregion of a region argument of the parent task;
furthermore, the subtask 
can only have permissions on the region that are compatible with the parent's 
permissions.   This requirement
plays an important role in our task scheduling algorithm (see
Section~\ref{sec:soop}).

%The declarations for the three subtasks are shown on lines 42-51.  
The  {\tt calc\_new\_currents} task reads and writes the wires subregion, 
the nodes subregion, and the ghost node region for its piece.
The {\tt distribute\_charge} subtask is a different phase that reads the piece's 
wires subregion and updates all nodes those wires touch.  However,
rather than requesting the ability to read and write the nodes (which would
require serialization of these tasks for correctness), the task declares that
it will use reorderable {\em reduction} operations and that the coherence requirement
is {\em atomic} rather than exclusive access. The final task is 
{\tt update\_voltages}, which writes the shared and private nodes for a piece
and reads the results of the previous task's reductions.

@@

\subsection{Data Types}
\label{subsec:datatypes}
The two basic data types in the circuit simulation are {\tt Node}s and
{\tt Wire}s declared on lines 1 and 2 of Listing~\ref{lst:code_ex}.
As is standard in region-based systems, allocating a {\tt Node}, {\tt
  Wire}, or any value in the heap requires naming the region
in which the value is allocated.  For example, {\tt new Node@r} returns
a reference to a new value of {\tt Node} type in region {\tt r}.  
Listing~\ref{lst:code_ex} illustrates one way of constructing partitioned
data structures in Legion: populate a region with data (using {\tt new}, note the allocation
of the data is omitted from the example) and then partition it
into subregions.  One can also first partition an empty region into subregions
and then allocate data in the subregions. We use both approaches in our
applications discussed in Section~\ref{sec:exp}.

Pointers to region values declare the region to which they point.  For
example, the definition of {\tt Wire} is parameterized on the region
{\tt rn} to which the {\tt Node} pointers in fields {\tt in\_nodes}
and {\tt out\_nodes} point.
% which guarantees that the two endpoints of
%a wire are always in the same region.

Regions are first-class values in Legion and can be stored in heap
data structures.  For example, a {\tt Circuit} (defined on line 3)
holds the regions for all nodes and wires of the circuit and a {\tt
  CircuitPiece} (defined on line 4) hold pointers to the private,
shared, and ghost node regions as well as the private wires regions of its piece of
the circuit.

%\include{part_fig}

%\include{code_ex}
%\include{part_fig}

>>>>

Listing~\ref{lst:code_ex} shows code for an electrical
circuit simulation, which takes a collection of
of wires and nodes where wires meet.  
The simulation repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.

The key decisions in a Legion program are how data is grouped into
regions and how those regions are {\em partitioned} into {\em
subregions}.  The goal is to pick an organization that makes explicit
in the algorithms which computations are independent.  A {\tt Circuit}
has two regions: undifferentiated collections of nodes and wires (line
3 of Listing~\ref{lst:code_ex}). An efficient parallel
implementation breaks this unstructured graph into pieces that can be
processed (mostly) independently. There is always some computation
across the wires between pieces, but an appropriate
region organization makes explicit which nodes and wires are
involved in inter-piece computation and which are involved in only
intra-piece computation.

% FIXME   We add here a description of how the colors print when printed greyscale.
%
Figure~\ref{sfig:part_fig:pvs} shows how the nodes
in a small graph might be split into three pieces.  Blue nodes are
attached by wires only to nodes in the same piece and are 
{\em private} to the piece.  Orange nodes on the boundary of a
piece are {\em shared} with (connected to) other pieces.
In the simulation, computations on the
private nodes of different pieces will be independent, while
computations on the shared nodes will require communication.  To make
this explicit in the program, we partition the nodes region into
private and shared subregions (line 18).  To partition a region, we
provide a {\em coloring}, which is a relation between the elements of
a region and a set of colors.  A partition is an object which given a
coloring and a region $r$ contains for each color $c$ a subregion of
$r$ containing all the elements of $r$ colored $c$. Note that the
partition into shared and private nodes is disjoint because each node has one color.

The private nodes and the shared nodes are partitioned again into the private and shared nodes for each circuit piece (lines 21
and 23);  both partitions are disjoint.  There is another useful partition of the shared nodes:
for a piece $i$, we will need the shared nodes that border $i$ in other pieces of the circuit.
This {\em ghost node} partition (line 25) has two interesting properties.  First, it is a second partition of the shared nodes:
we have two views on to the same collection of data.  Second, the ghost node partition is {\em aliased}, meaning the subregions
are not disjoint: a node may border several different circuit pieces and belong to more than one ghost node subregion (thus, {\tt node\_neighbor\_map} on line 25 assigns more than one color to some nodes).
The private, shared, and ghost node subregions
for the upper-left patch of the example graph are shown in
Figures~\ref{sfig:part_fig:p_i}, \ref{sfig:part_fig:s_i}, and
\ref{sfig:part_fig:g_i} respectively.

Figure~\ref{sfig:part_fig:tree} shows the final hierarchy of node
partitions and subregions. The $*$ symbol indicates a partition is disjoint. This
{\em region tree} data structure plays an important role in scheduling
tasks for out-of-order execution (see Section~\ref{sec:soop}).
The organization of the wires is much simpler: a single disjoint partition
that assigns each wire to one piece (line 16).

%
% FIXME
% Sean wants to use the term task through this paragraph; however we haven't defined it yet.
% I think it is OK as is since the term is introduced at the end of the paragraph.
%
Line 9 declares the main simulator function, which takes a {\tt
Circuit} to be simulated.  This function specifies the regions it 
accesses and the privileges and coherence it requires of those regions.
In this case, the {\tt RWE} annotation specifies that the function
accesses the regions {\tt c.r\_all\_nodes} and {\tt c.r\_all\_wires}
with read-write privileges and {\em exclusive} coherence (i.e., no other
function can access these two regions concurrently).  The simulation 
reads and writes all nodes and wires, and it must be done
with exclusive access to ensure correct results.  Privileges specify what
the function can do with the regions; coherence specifies what other
functions can do with the regions concurrently.  Functions that
declare their accessed regions, privileges, and coherence are called {\em tasks}
and are the unit of parallel execution in Legion.

Lines 31-57 form the bulk of the actual simulation, performing three
passes over the circuit in each iteration.  Each pass loops over an
array of pieces (constructed on lines 27-30 from the partitions),
spawning a task for each piece.  There are no explicit requests for
parallel execution ({\tt spawn} simply indicates a task call and does
not require parallel execution) nor is there synchronization required
between the passes.  Both the fact that the tasks can be run in
parallel within a pass and the required inter-pass dependencies are
determined automatically by the Legion runtime based on the region
access annotations on the task declarations.  The tasks spawned on
lines 32-34 are {\em subtasks} of the main {\tt simulate\_circuit}
task. A subtask can only take a region argument that is a region
argument of the parent task or a subregion of a region argument of the
parent task; furthermore, the subtask can only have permissions on the
region that are compatible with the parent's permissions.  
%This
%requirement plays an important role in our task scheduling algorithm
%(see Section~\ref{sec:soop}).

The  {\tt calc\_new\_currents} task reads and writes the wires subregion 
and reads the private, shared, and ghost node regions for its piece.
The {\tt distribute\_charge} subtask is a different phase that reads the piece's 
wires subregion and updates all nodes those wires touch.  However,
rather than requesting the ability to read and write the nodes (which would
require serialization of these tasks for correctness), the task declares that
it will use reorderable {\em reduction} operations and that the coherence requirement
is {\em atomic} rather than exclusive access. The final task is 
{\tt update\_voltages}, which writes the shared and private nodes for a piece
and reads the results of the previous task's reductions.

@@

%\subsection{Data Types}
%\label{subsec:datatypes}

The two basic data types in the circuit simulation are {\tt Node}s and
{\tt Wire}s declared on lines 1 and 2 of Listing~\ref{lst:code_ex}.
As is standard in region-based systems, allocating a {\tt Node}, {\tt
  Wire}, or any value in the heap requires naming the region
in which the value is allocated.  For example, {\tt new Node@r} returns
a reference to a new value of {\tt Node} type in region {\tt r}.  
Listing~\ref{lst:code_ex} illustrates one way of constructing partitioned
data structures in Legion: populate a region with data (using {\tt new}, note the example
makes use of whatever data has been allocated by the caller of {\tt simulate\_circuit})
and then partition it into subregions.  One can also first partition an empty region into subregions
and then allocate data in the subregions. We use both approaches in our
applications (see Section~\ref{sec:exp}).

Pointers to region values declare the region to which they point.  For
example, the definition of {\tt Wire} is parameterized on the region
{\tt rn} to which the {\tt Node} pointers in fields {\tt in\_nodes}
and {\tt out\_nodes} point.
% which guarantees that the two endpoints of
%a wire are always in the same region.
Regions are first-class values in Legion and can be stored in heap
data structures.  For example, a {\tt Circuit} (defined on line 3)
holds the regions for all nodes and wires of the circuit and a {\tt
  CircuitPiece} (defined on line 4) hold pointers to the private,
shared, and ghost node regions as well as the private wires regions of its piece of
the circuit.

### 2012-05-04 doc/sc2012/legion.tex
revision

\begin{abstract}
Modern parallel architectures have both heterogeneous processors and
deep, complex memory hierarchies.  We present Legion, a programming
model and runtime system for achieving high performance on these
machines.  Legion is organized around {\em logical regions}, which
express both locality and independence of program data, and {\em
  tasks}, functions that perform computations on regions.  We describe
a runtime system that dynamically extracts parallelism from
Legion programs, using a distributed, parallel scheduling algorithm
that identifies both independent tasks and nested parallelism.  Legion
also enables explicit, programmer controlled movement of data through
the memory hierarchy and placement of tasks based on locality
information via a novel mapping interface.  We evaluate our Legion
implementation using four applications: fluid-flow on a regular grid,
a three-level AMR code solving a standard heat diffusion equation, a circuit simulation,
and LINPACK.
%Running on a 4 node cluster with 8 total GPUs 
%and 4 levels of memory hierarchy, our implementation of Legion 
%achieves up to 5.9X speedup over a single CPU-GPU node on real-world applications.
\end{abstract}

@@

Modern parallel machines are increasingly complex, with deep,
distributed memory hierarchies and heterogeneous processing units.  Because the
costs of communication within these architectures vary by several orders of magnitude, 
the penalty for mistakes in the placement of data or computation is usually very poor performance.  Thus,
to achieve good performance it is crucial that the programmer
and the compiler be able to reason about {\em locality} (data resident
close to computation that uses it) and {\em independence} (computations
operating on disjoint data, and therefore not requiring communication and able to be placed in possibly distant parts of the machine).  
Most contemporary
programming systems have no facilities for the programmer to express locality
and independence.  The few languages that do focus primarily on
array-based locality \cite{Fatahalian06,CHAPEL04,UPC99} and 
avoid irregular pointer data structures.
%encode such structures in arrays, which 
%data structures encoded in arrays which are impossible for compilers
%and runtime systems to reason about with regard to locality and independence.

In this paper we describe Legion, a parallel programming system based
on using {\em logical regions} to describe the organization of data.
A logical region names a set of objects.  Logical regions are first
class values in Legion, and in particular may be passed as arguments
to distinguished functions called {\em tasks} that access the data in
those regions, providing locality information.  Logical regions may be
{\em partitioned} into (usually) disjoint {\em subregions}, providing
information for determining independence of computations.
Furthermore, computations access logical regions with particular {\em
  privileges} (e.g., {\em read-only} and {\em read-write}, among
others) and {\em coherence} (e.g., {\em exclusive access} and {\em
  atomic access}, among others).  Privileges express how a task may use
its region arguments, providing data dependence information that
is used to guide the extraction of parallelism.  For example, if two tasks
access the same region with read-only privileges the two tasks
can potentially be run in parallel.
Coherence properties express the required semantics of concurrent
accesses to the region.  For example, if the program executes $f_1(r); f_2(r)$ and tasks $f_1$
and $f_2$ both require {\em exclusive} access to region $r$, then Legion guarantees that 
$f_1(r)$ completes before $f_2(r)$ begins.   On the other hand, if
the tasks access $r$ with {\em atomic} coherence, then Legion
guarantees only atomicity of the tasks with respect to $r$: either task $f_1(r)$
appears to run entirely before $f_2(r)$ or vice versa.

@@

To introduce the programming model, we begin with an extended example
of a circuit simulation in
Section~\ref{sec:ex}. This example illustrates regions, tasks,
permissions and coherence properties, the interactions between them,
and how these building blocks are assembled into a Legion
program.  Subsequent sections each describe a contribution in the
implementation and evaluation of Legion:

@@

\item We define a {\em software out-of-order processor}, or SOOP, for
  scheduling tasks with region arguments in a manner analogous to how
  out-of-order hardware schedulers process instructions with register
  arguments (Section~\ref{sec:soop}).  In addition to pipelining the execution of tasks over
  several stages, our SOOP is concurrent and distributed across the machine and is
  also hierarchical and naturally extracts nested
  parallelism (because tasks may recursively spawn subtasks).

@@

\item A central part of orchestrating a program's execution is how
tasks are assigned (or {\em mapped}) to processors and how physical instances of logical regions are
mapped to specific memory units within the machine  (Section~\ref{sec:mapping}).  Often there is
application-specific information that, if exploited, results in better
mappings than a generic, one-size-fits-all mapping strategy.  We describe
a {\em mapping interface} that allows programmers to give the SOOP a specification
of how to map tasks and regions for a specific application, or even part of an
application.  This mapping API is carefully defined so that
any user-supplied mapping strategy can only affect the performance of applications,
not their correctness.  
%Thus, user mappings only affect the time and resources required
%to run the program and cannot change the program's output.

\item We present the results of experiments on four applications:
  fluid-flow on a regular grid, a three-level AMR code solving a
  standard heat diffusion equation, a circuit simulation, and LINPACK (Section~\ref{sec:exp}).
  We run each of these four codes and best available reference
  versions on three different clusters of multicore processors with
  attached GPUs: a small cluster of our own (4 nodes with 2 6-core CPUs and 2 GPUs per node), a medium-sized cluster
  designed for visualization applications (10 nodes with 2 6-core CPUs and 5 GPUs per node), and up to a third of the
  Keeneland supercomputer (32 nodes, 2 6-core CPUs and 3 GPUs per node)

>>>>

\begin{abstract}
Modern parallel architectures have both heterogeneous processors and
deep, complex memory hierarchies.  We present Legion, a programming
model and runtime system for achieving high performance on these
machines.  Legion is organized around {\em logical regions}, which
express both locality and independence of program data, and {\em
  tasks}, functions that perform computations on regions.  We describe
a runtime system that dynamically extracts parallelism from
Legion programs, using a distributed, parallel scheduling algorithm
that identifies both independent tasks and nested parallelism.  Legion
also enables explicit, programmer controlled movement of data through
the memory hierarchy and placement of tasks based on locality
information via a novel mapping interface.  We evaluate our Legion
implementation on three applications: fluid-flow on a regular grid,
a three-level AMR code solving a heat diffusion equation, and 
a circuit simulation.
%
% FIXME: put some performance numbers here
%
\end{abstract}

@@

Modern parallel machines are increasingly complex, with deep,
distributed memory hierarchies and heterogeneous processing units.  Because the
costs of communication within these architectures vary by several orders of magnitude, 
the penalty for mistakes in the placement of data or computation is usually very poor performance.  Thus,
to achieve good performance it is crucial that the programmer
and the compiler be able to reason about {\em locality} (data resident
close to computation that uses it) and {\em independence} (computations
operating on disjoint data, and therefore not requiring communication and able to be placed in possibly distant parts of the machine).  
Most contemporary
programming systems have no facilities for the programmer to express locality
and independence.  The few languages that do focus primarily on
array-based locality \cite{Fatahalian06,CHAPEL04,UPC99} and 
avoid irregular pointer data structures.  
%
% FIXME
% Sean would like a bit more on why other systems don't support pointer data structures (e.g., the difficulty of
% alias analysis.  I thought about it and feel it is too much of a digression: mentioning that in the first paragraph
% makes it sound like it will be an important part of the paper.
%
%encode such structures in arrays, which 
%data structures encoded in arrays which are impossible for compilers
%and runtime systems to reason about with regard to locality and independence.

In this paper we describe Legion, a parallel programming system based
on using {\em logical regions} to describe the organization of data
and to make explicit relationships useful for reasoning about locality
and independence.
A logical region names a set of objects.  Logical regions are first-class 
values in Legion and may be dynamically allocated, deleted and stored
in data structures. Regions can also be passed as arguments
to distinguished functions called {\em tasks} that access the data in
those regions, providing locality information.  Logical regions may be
{\em partitioned} into (usually) disjoint {\em subregions}, providing
information for determining independence of computations.
Furthermore, computations access logical regions with particular {\em
  privileges} ({\em read-only}, {\em read-write}, and {\em reduce}) and {\em coherence} (e.g., {\em exclusive access} and {\em
  atomic access}, among others).  Privileges express how a task may use
its region arguments, providing data dependence information that
is used to guide the extraction of parallelism.  For example, if two tasks
access the same region with read-only privileges the two tasks
can potentially be run in parallel.
Coherence properties express the required semantics of concurrent region accesses.
For example, if the program executes $f_1(r); f_2(r)$ and tasks $f_1$
and $f_2$ both require {\em exclusive} access to region $r$, then Legion guarantees the result will
be as if $f_1(r)$ completes before $f_2(r)$ begins.   On the other hand, if
the tasks access $r$ with {\em atomic} coherence, then Legion
guarantees only atomicity of the tasks with respect to $r$: either task $f_1(r)$
appears to run entirely before $f_2(r)$ or vice versa.

@@

To introduce the programming model, we present
a circuit simulation in
Section~\ref{sec:ex}, illustrating regions, tasks,
permissions and coherence properties, the interactions between them,
and how these building blocks are assembled into a Legion
program.  Subsequent sections each describe a contribution in the
implementation and evaluation of Legion:

@@

\item We define a {\em software out-of-order processor}, or SOOP, for
  scheduling tasks with region arguments in a manner analogous to how
  out-of-order hardware schedulers process instructions with register
  arguments (Section~\ref{sec:soop}).  In addition to pipelining the execution of tasks over
  several stages, our SOOP is distributed across the machine and is
  also hierarchical and naturally extracts nested
  parallelism (because tasks may recursively spawn subtasks).

@@

\item Of central importance is how
tasks are assigned (or {\em mapped}) to processors and how physical instances of logical regions are
mapped to specific memory units (Section~\ref{sec:mapping}).  Often using
application-specific information results in better
mappings than a generic mapping strategy.  We describe
a {\em mapping interface} that allows programmers to give the SOOP a specification
of how to map tasks and regions for a specific application, or even part of an
application.  This mapping API is designed so that
any user-supplied mapping strategy can only affect the performance of applications,
not their correctness.  
%Thus, user mappings only affect the time and resources required
%to run the program and cannot change the program's output.

\item We present the results of experiments on three applications:
  fluid-flow on a regular grid, a three-level AMR code solving a
  standard heat diffusion equation and a circuit simulation.
  We run each of these four codes and best available reference
  versions on three different clusters of multicore processors with
  attached GPUs: a small cluster of our own (4 nodes with 12 CPUs and 2 GPUs per node), a medium-sized cluster
  designed for visualization applications (10 nodes with 12 CPUs and 5 GPUs per node), and up to a third of the
  Keeneland supercomputer (32 nodes, 12 CPUs and 3 GPUs per node).
%
% FIXME Give the same numbers as in the abstract.
%
